Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-29 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd
@NidhiVinod , please verify that this is correct.

*beta carotene and lutein are part of the caratenoids pool that responsible for light absorption, and also heat energy dissipation*

Biochemical protection against foliage light and heat damage also increases with irradiance, and thus tends to be higher in the upper canopy than in the understory (Table `r table_leaf_traits`). 
Carotenoids and xanthophyll cycle pigments -- violaxanthin, antheraxanthin and zeaxanthin (VAZ) -- can play a role in antioxidant scavenging and converting excess excitation energy into heat to reduce photoinhibition [@niinemetsPhotosynthesisResourceDistribution2007a; @mathurPhotosyntheticEfficiencySun2018a; @niinemetsAcclimationHighIrradiance1998]. 
The dissipation of excess light energy is also important for leaf acclimation to higher temperatures if high $T_{leaf}$ impairs the photochemical energy dissipation pathway [@havauxTemperaturedependentAdjustmentThermal1996]. 
Capacity for heat-sensitive emissions of volatile organic compounds (VOCs), including isoprenes and monoterpenes, enhances photosynthetic thermal tolerance by regulating antioxidant defenses and other metabolic processes [@sharkeyIsopreneEmissionPlants2008; @taylorCapacityEmitIsoprene2019; @monsonLeafIsopreneEmission2021; **REF ON MONOTERPENES?**]. 
VOC emission capacity is a species-level trait [**REFS**], such that intrapspecific variation across across vertical gradients are not applicable, but frequency of this trait varies across the vertical gradient (see section 3.2), as do emission rates [see section 4.4, @taylorNewFieldInstrument2021].

>>>>

Biochemical protection against foliage light and heat damage also increases with irradiance, and thus tends to be higher in the upper canopy than in the understory (Table `r table_leaf_traits`). 
Carotenoids (e.g., beta carotene and lutein), including xanthophyll cycle pigments (i.e., violaxanthin, antheraxanthin and zeaxanthin, VAZ), can play a role in antioxidant scavenging and converting excess excitation energy into heat to reduce photoinhibition [@niinemetsPhotosynthesisResourceDistribution2007a; @mathurPhotosyntheticEfficiencySun2018a; @niinemetsAcclimationHighIrradiance1998]. 
The dissipation of excess light energy is also important for leaf acclimation to higher temperatures if high $T_{leaf}$ impairs the photochemical energy dissipation pathway [@havauxTemperaturedependentAdjustmentThermal1996]. 
Capacity for heat-sensitive emissions of volatile organic compounds (VOCs), including isoprenes and monoterpenes, enhances photosynthetic thermal tolerance by regulating antioxidant defenses and other metabolic processes [@sharkeyIsopreneEmissionPlants2008; @taylorCapacityEmitIsoprene2019; @monsonLeafIsopreneEmission2021; **REF ON MONOTERPENES?**]. 
VOC emission capacity is a species-level trait [**REFS**], such that intrapspecific variation across across vertical gradients are not applicable, but frequency of this trait varies across the vertical gradient (see section 3.2), as do emission rates [see section 4.4, @taylorNewFieldInstrument2021].
